---
title: "ACTT validation study"
format: 
  html:
    self-contained: true
editor_options: 
  chunk_output_type: inline
---

# 📦 Load Required Libraries

```{r}
#| warning: false

# Data manipulation and visualisation
library(tidyverse)
# Classification metrics (confusion matrix, sensitivity, specificity)
library(caret)
# Multi-model inference (available for R² if needed)
library(MuMIn)
# Colour palettes inspired by the Metropolitan Museum of Art
library(MetBrewer)
# Tidy model output (used for Tukey HSD results)
library(broom)
# Multi-panel figure composition
library(cowplot)
# ROC curves and AUC
library(pROC)
```

# 🎨 Set Theme and Color Palette

```{r}
# Select 4 colours from the 11-colour "Archambault" palette.
# These map to GT3X+ (hip), ACTT (hip), ACTT (wrist), GT3X+ (wrist)
# throughout all figures.
pal <- met.brewer("Archambault", 11)[c(1, 4, 6, 8)]

# Apply a consistent ggplot2 theme across all figures
theme_set(theme_bw() + 
            theme_bw(base_size = 14, 
                     base_family = "Helvetica") + 
            theme(panel.border = element_rect(linewidth = 1.5)))
```

# 📁 Load and View Data

```{r}
#| warning: false

# Single CSV with one row per participant × condition.
# Columns: ID, Conditions, METs, BMI, GT3X+ (hip), GT3X+ (wrist),
#          ACTT (hip), ACTT (wrist)
# NOTE: The 'Sit-Stands' condition is in the data but excluded from
#       all primary analyses below.
df <- readr::read_csv(file = 'data/acttrust_validation_full.csv') %>% 
  dplyr::mutate(Sex = dplyr::recode(Sex, '1' = 'Female', '2' = 'Male'))
df
```

# 🔍 ANOVA and Post-Hoc Tests

## ⚖️️ ANOVA by Device and Condition

```{r}
# Reshape to long format so that all four device-placement columns
# become a single 'act' factor. Sit-Stands excluded.
# Square-root transform applied to stabilise variance.
# Reference levels: GT3X+ (hip) for device; Rest for condition.
# Model: sqrt(act_value) ~ Conditions:act (interaction only, no main effects)
df %>% 
  dplyr::filter(Conditions != 'Sit-Stands') %>% 
  tidyr::pivot_longer(cols = contains(c('GT', 'ACT')), 
                      names_to = 'act', 
                      values_to = 'act_value') %>% 
  dplyr::mutate(act = relevel(as.factor(act), ref = 'GT3X+ (hip)')) %>% 
  dplyr::mutate(Conditions = relevel(as.factor(Conditions), ref = 'Rest')) %>% 
  lm(formula = sqrt(act_value) ~ Conditions:act, data = .) %>% 
  aov() %>% 
  summary()
```

## 🔬 Tukey Post-Hoc

```{r}
# Re-fit the ANOVA model and apply Tukey's Honestly Significant Difference
# test for all pairwise comparisons between Conditions × device combinations.
# Results are tidied into a data frame using broom::tidy().
aov_model <- df %>% 
  dplyr::filter(Conditions != 'Sit-Stands') %>% 
  tidyr::pivot_longer(cols = contains(c('GT', 'ACT')), 
                      names_to = 'act', 
                      values_to = 'act_value') %>% 
  dplyr::mutate(act = relevel(as.factor(act), ref = 'GT3X+ (hip)')) %>% 
  dplyr::mutate(Conditions = relevel(as.factor(Conditions), ref = 'Rest')) %>% 
  lm(formula = sqrt(act_value) ~ Conditions:act, data = .) %>% 
  aov()

tukey_res <- TukeyHSD(aov_model)
tidy_tukey <- tidy(tukey_res)

tidy_tukey
```

# 🔗 Correlation Matrix of Devices

```{r}
# Compute Pearson correlations between the four device-placement combinations
# (Sit-Stands excluded; NAs dropped).
# Visualised as a heatmap (Figure 2a).
cor_matrix <- df %>% 
  dplyr::filter(Conditions != 'Sit-Stands') %>% 
  dplyr::select(contains(c('GT', 'ACT'))) %>% 
  tidyr::drop_na() %>% 
  cor(., method = "pearson")

# Melt to long format for ggplot2; reverse Var2 axis for conventional layout
cor_melt <- reshape2::melt(cor_matrix)
cor_melt$Var2 <- factor(cor_melt$Var2, levels = rev(levels(cor_melt$Var2)))

ggplot(cor_melt, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient(low  = met.brewer("Archambault", 11)[6], 
                      high = met.brewer("Archambault", 11)[4], 
                      name = "Pearson\nCorrelation\n", 
                      labels = scales::label_number(accuracy = 0.01)) +
  xlab(element_blank()) + 
  ylab(element_blank()) + 
  theme(axis.text.x     = element_text(angle = 45, vjust = 1, size = 12, hjust = 1), 
        axis.text.y     = element_text(size = 12),
        legend.title    = element_text(size = 14), 
        legend.text     = element_text(size = 12)) +
  coord_fixed()
```

# 📈 Linear Model and Predictions

```{r}
# Reshape data for modelling: long format, Sit-Stands excluded, NAs dropped.
# 'act' factor is relevelled so that GT3X+ (hip) is the reference level —
# all other device coefficients are contrasts against this baseline.
dfm <- df %>% 
  dplyr::filter(!Conditions %in% c('Sit-Stands')) %>% 
  dplyr::mutate(Conditions = factor(Conditions)) %>%
  tidyr::drop_na() %>%
  tidyr::pivot_longer(cols = contains(c('GT', 'ACT')), 
                      names_to  = 'act', 
                      values_to = 'act_value') %>% 
  dplyr::mutate(act = relevel(as.factor(act), ref = 'GT3X+ (hip)')) 

# General linear model (Equation 1 in manuscript):
#   MET^0.5 ~ Activity^0.5 * device_placement
# The * operator includes main effects and the interaction, allowing each
# device to have its own slope and intercept (Equation 2).
lm(formula = sqrt(METs) ~ sqrt(act_value) * act, data = dfm) %>% 
  summary()
```

# 📊 Plot Regression Fits

```{r}
# Extract model coefficients for plotting device-specific regression lines.
# cf[1]        = intercept (GT3X+ hip reference)
# cf[2]        = slope     (GT3X+ hip reference)
# cf[3:5]      = intercept adjustments for ACTT(hip), ACTT(wrist), GT3X+(wrist)
# cf[6:8]      = slope adjustments for the same devices
cf <- lm(formula = sqrt(METs) ~ sqrt(act_value) * act, data = dfm) %>% 
  coef()

dfm %>% 
  ggplot2::ggplot(aes(x = sqrt(act_value), y = sqrt(METs))) +
  geom_point(aes(colour = act, shape = Conditions), size = 2.5) + 
  # GT3X+ (hip) — reference line
  geom_abline(intercept = cf[1],           slope = cf[2],           colour = pal[1], linewidth = 1.5) +
  # ACTT (hip)
  geom_abline(intercept = cf[1] + cf[3],   slope = cf[2] + cf[6],   colour = pal[2], linewidth = 1.5) +
  # ACTT (wrist)
  geom_abline(intercept = cf[1] + cf[4],   slope = cf[2] + cf[7],   colour = pal[3], linewidth = 1.5) +
  # GT3X+ (wrist)
  geom_abline(intercept = cf[1] + cf[5],   slope = cf[2] + cf[8],   colour = pal[4], linewidth = 1.5) +
  scale_colour_manual(values = pal) + 
  guides(colour = guide_legend(override.aes = list(size = 5)), 
         shape  = guide_legend(override.aes = list(size = 5))) + 
  theme(legend.background = element_blank(), 
        legend.title       = element_blank()) + 
  xlab(expression(Activity^{1/2})) + 
  ylab(expression(MET^{1/2}))

# Print full model summary for reporting (Table 2 equations derived from cf[])
lm(formula = sqrt(METs) ~ sqrt(act_value) * act, data = dfm) %>% 
  summary()
```

# 🔮 Classification Metrics

## 🤖 Fit and Classify

```{r}
# Obtain in-sample fitted METs by squaring the model's fitted sqrt(METs) values
fit_MET <- lm(formula = sqrt(METs) ~ sqrt(act_value) * act, 
              data = dfm) %>% 
  fitted()

fit_MET <- fit_MET^2
```

```{r}
lm(formula = sqrt(METs) ~ sqrt(act_value) * act, 
              data = dfm) %>% 
  summary()
```

```{r}
# Helper function: classify MET values into four PA intensity bands
#   [0, 3)   = Light
#   [3, 6)   = Moderate
#   [6, 9)   = Vigorous
#   [9, Inf) = Very Vigorous
calc_met_class <- function(met_values) {
  classes <- cut(met_values,
                 breaks = c(-Inf, 3, 6, 9, Inf),
                 labels = c('[0,3)', '[3,6)', '[6,9)', '[9,inf)'),
                 right  = FALSE)
  return(classes)
}

# Build prediction data frame and classify both actual and predicted METs
dft_pred <- dfm %>% 
  dplyr::select(Conditions, METs, act_value, BMI, ID) %>%
  dplyr::mutate(fit_MET^2) %>% 
  dplyr::mutate(actual_class    = cut(METs,
                                      breaks = c(-Inf, 3, 6, 9, Inf),
                                      labels = c('[0,3)', '[3,6)', '[6,9)', '[9,inf)'),
                                      right  = FALSE)) %>% 
  dplyr::mutate(predicted_class = calc_met_class(fit_MET))

# Aggregate confusion matrix (all devices pooled) — reproduces Figure 2c and Table 3
caret::confusionMatrix(dft_pred$predicted_class, dft_pred$actual_class)
cm    <- caret::confusionMatrix(dft_pred$predicted_class, dft_pred$actual_class)

# Normalise by row (true class) for heatmap display
cm_df <- as.data.frame(prop.table(cm$table, margin = 1))
cm_df$Reference <- factor(cm_df$Reference, levels = rev(levels(cm_df$Reference)))
```

## 📊 Confusion Matrix Plot

```{r}
# Heatmap of row-normalised confusion matrix (Figure 2c)
ggplot(data = cm_df, aes(x = Prediction, y = Reference, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.2f", Freq)), color = "white", size = 5) +
  scale_fill_gradient(low = "white", high = met.brewer("Archambault", 11)[3]) +
  theme(legend.position = "none") + 
  labs(title = "Confusion Matrix", x = "Predicted Class", y = "True Class")
```

## 📈 AUC by Class

```{r}
# One-vs-rest AUC for each intensity class.
# fit_MET (continuous) is used as the predictor score.
actual_class <- factor(dft_pred$actual_class, 
                       levels = c('[0,3)', '[3,6)', '[6,9)', '[9,inf)'))

auc_results <- data.frame(Class = character(), AUC = numeric())

for (lvl in levels(actual_class)) {
  binary_actual <- as.numeric(actual_class == lvl)  # 1 = true class, 0 = all others
  roc_obj       <- roc(binary_actual, fit_MET)
  auc_val       <- auc(roc_obj)
  auc_results   <- rbind(auc_results, data.frame(Class = lvl, AUC = auc_val))
}

print(auc_results)
```

# 🧮 Device-Specific Cutoffs

```{r}
#| warning: false

# Invert the regression equation to find the activity count threshold
# corresponding to target MET values (3, 6, 9, and additionally 4 and 8).
#
# From MET^0.5 = β†0 + β†1 * Activity^0.5 :
#   Activity = ((MET^0.5 - β†0) / β†1)^2
#
# β†0 and β†1 are device-specific intercepts and slopes derived from cf[].
cutoffs <- tibble::tibble(MET_values = c(3, 6, 9, 4, 8)) %>% 
  dplyr::mutate(
    `GT3X+ (hip)`   = (((sqrt(MET_values) - cf[1])         /  cf[2]        ))^2, 
    `ACTT (hip)`    = (( sqrt(MET_values) - cf[1] - cf[3]) / (cf[2] + cf[6]))^2, 
    `ACTT (wrist)`  = (( sqrt(MET_values) - cf[1] - cf[4]) / (cf[2] + cf[7]))^2, 
    `GT3X+ (wrist)` = (( sqrt(MET_values) - cf[1] - cf[5]) / (cf[2] + cf[8]))^2
  ) %>% 
  t()

# Display as formatted table (Table 2 in manuscript)
cutoffs[2:5, ] %>% 
  tibble::as_tibble() %>% 
  dplyr::mutate(Actimeter = c('GT3X+ (hip)', 'ACTT (hip)', 'ACTT (wrist)', 'GT3X+ (wrist)')) %>% 
  dplyr::rename('3' = V1, '6' = V2, '9' = V3, '4' = V4, '8' = V5) %>%
  gt::gt()
```

# 📊 Summary Plots

```{r}
#| warning: false
#| fig-height: 12
#| fig-width: 8

# Figure 1: two-panel composite
#   Panel a) METs by treadmill condition
#   Panel b) Activity counts by condition, faceted by device × placement

sum_plot <- cowplot::plot_grid(
  
  # Panel a — METs
  df %>% 
    tidyr::pivot_longer(cols = contains(c('GT', 'ACT', 'MET')), 
                        names_to  = 'act', 
                        values_to = 'act_value') %>% 
    dplyr::filter(act == 'METs') %>% 
    dplyr::filter(Conditions != 'Sit-Stands') %>% 
    dplyr::mutate(Conditions = relevel(as.factor(Conditions), ref = 'Rest')) %>% 
    ggplot2::ggplot(aes(x = Conditions, y = act_value)) + 
    geom_point(aes(colour = act), 
               position = position_jitterdodge(0.5), 
               size = 3, alpha = 0.4) + 
    geom_boxplot(fill = NA, outliers = FALSE) + 
    geom_smooth(aes(group = act), 
                colour   = 'grey50', 
                linetype = 'dashed', 
                method   = 'loess', 
                se       = FALSE, 
                size     = 0.8) + 
    facet_wrap(~ act, scales = 'free_y') + 
    scale_colour_manual(values = 'grey70') + 
    theme(legend.position  = "none", 
          axis.text.x      = element_text(angle = 45, hjust = 1)) + 
    ylab(expression("Metabolic Equivalents (VO"^2*")")) + 
    xlab(element_blank()), 
  
  # Panel b — Activity counts, faceted by device × placement
  df %>% 
    tidyr::pivot_longer(cols = contains(c('GT', 'ACT', 'MET')), 
                        names_to  = 'act', 
                        values_to = 'act_value') %>% 
    dplyr::filter(act != 'METs') %>% 
    dplyr::filter(Conditions != 'Sit-Stands') %>% 
    dplyr::mutate(
      act        = relevel(as.factor(act), ref = 'GT3X+ (hip)'), 
      Conditions = relevel(as.factor(Conditions), ref = 'Rest'),
      act        = factor(act, levels = c("GT3X+ (hip)", "GT3X+ (wrist)", 
                                          "ACTT (hip)",  "ACTT (wrist)"))
    ) %>% 
    ggplot2::ggplot(aes(x = Conditions, y = act_value)) + 
    geom_point(aes(colour = act), 
               position = position_jitterdodge(1.2), 
               size = 2, alpha = 0.4) + 
    geom_boxplot(fill = NA, outliers = FALSE) + 
    geom_smooth(aes(group = act), 
                colour   = 'grey50', 
                linetype = 'dashed', 
                method   = 'loess', 
                se       = FALSE, 
                size     = 0.8) + 
    facet_wrap(~ act, scales = 'free_y') + 
    scale_colour_manual(values = c(pal[1], pal[4], pal[2], pal[3])) + 
    theme(legend.position = "none", 
          axis.text.x     = element_text(angle = 45, hjust = 1)) + 
    ylab(expression("Activity Counts (counts·min"^-1*")")) + 
    xlab(element_blank()), 
  
  ncol   = 1, 
  labels = c('a', 'b')
)

sum_plot
```

# Device-specific confusion matrices

```{r}
# Re-fit the model and append fitted METs to dfm for device-level filtering
fit_MET <- lm(formula = sqrt(METs) ~ sqrt(act_value) * act, data = dfm) %>%
  fitted() %>%
  `^`(2)

dfm <- dfm %>%
  mutate(fit_MET = fit_MET)

# --- GT3X+ (hip) -------------------------------------------------------------
dft_hip <- dfm %>%
  filter(act == 'GT3X+ (hip)') %>%
  select(Conditions, METs, act_value, BMI, ID, fit_MET) %>%
  mutate(
    actual_class    = cut(METs, breaks = c(-Inf, 3, 6, 9, Inf),
                          labels = c('[0,3)', '[3,6)', '[6,9)', '[9,inf)'), 
                          right  = FALSE),
    predicted_class = calc_met_class(fit_MET)
  )

cm_hip    <- caret::confusionMatrix(dft_hip$predicted_class, dft_hip$actual_class)
cm_df_hip <- as.data.frame(prop.table(cm_hip$table, margin = 1))
cm_df_hip$Reference <- factor(cm_df_hip$Reference, levels = rev(levels(cm_df_hip$Reference)))

plot_hip <- ggplot(data = cm_df_hip, aes(x = Prediction, 
                                         y = Reference, 
                                         fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.2f", Freq)), 
            color = "white", 
            size = 5) +
  scale_fill_gradient(low = "white", 
                      high = pal[1]) +
  theme(legend.position = "none") +
  labs(title = "GT3X+ (hip)", 
       x = "Predicted Class", 
       y = "True Class")

# --- GT3X+ (wrist) -----------------------------------------------------------
dft_wrist <- dfm %>%
  filter(act == 'GT3X+ (wrist)') %>%
  select(Conditions, METs, act_value, BMI, ID, fit_MET) %>%
  mutate(
    actual_class    = cut(METs, breaks = c(-Inf, 3, 6, 9, Inf),
                          labels = c('[0,3)', '[3,6)', '[6,9)', '[9,inf)'), 
                          right  = FALSE),
    predicted_class = calc_met_class(fit_MET)
  )

cm_wrist <- caret::confusionMatrix(dft_wrist$predicted_class, 
                                      dft_wrist$actual_class)
cm_df_wrist <- as.data.frame(prop.table(cm_wrist$table, 
                                        margin = 1))
cm_df_wrist$Reference <- factor(cm_df_wrist$Reference, 
                                levels = rev(levels(cm_df_wrist$Reference)))

plot_wrist <- ggplot(data = cm_df_wrist, 
                     aes(x = Prediction, 
                         y = Reference, 
                         fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.2f", Freq)), 
            color = "white", 
            size = 5) +
  scale_fill_gradient(low = "white", 
                      high = pal[4]) +
  theme(legend.position = "none") +
  labs(title = "GT3X+ (wrist)", 
       x = "Predicted Class", 
       y = "True Class")

# --- ACTT (hip) --------------------------------------------------------------
dft_actt_hip <- dfm %>%
  filter(act == 'ACTT (hip)') %>%
  select(Conditions, METs, act_value, BMI, ID, fit_MET) %>%
  mutate(
    actual_class    = cut(METs, breaks = c(-Inf, 3, 6, 9, Inf),
                          labels = c('[0,3)', '[3,6)', '[6,9)', '[9,inf)'), 
                          right  = FALSE),
    predicted_class = calc_met_class(fit_MET)
  )

cm_actt_hip    <- caret::confusionMatrix(dft_actt_hip$predicted_class, 
                                         dft_actt_hip$actual_class)
cm_df_actt_hip <- as.data.frame(prop.table(cm_actt_hip$table, margin = 1))
cm_df_actt_hip$Reference <- factor(cm_df_actt_hip$Reference, 
                                   levels = rev(levels(cm_df_actt_hip$Reference)))

plot_actt_hip <- ggplot(data = cm_df_actt_hip, aes(x = Prediction, y = Reference, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.2f", Freq)), color = "white", size = 5) +
  scale_fill_gradient(low = "white", high = pal[2]) +
  theme(legend.position = "none") +
  labs(title = "ACTT (hip)", x = "Predicted Class", y = "True Class")

# --- ACTT (wrist) ------------------------------------------------------------
dft_actt_wrist <- dfm %>%
  filter(act == 'ACTT (wrist)') %>%
  select(Conditions, METs, act_value, BMI, ID, fit_MET) %>%
  mutate(
    actual_class    = cut(METs, 
                          breaks = c(-Inf, 3, 6, 9, Inf),
                          labels = c('[0,3)', '[3,6)', '[6,9)', '[9,inf)'), 
                          right  = FALSE),
    predicted_class = calc_met_class(fit_MET)
  )

cm_actt_wrist <- caret::confusionMatrix(dft_actt_wrist$predicted_class, 
                                        dft_actt_wrist$actual_class)
cm_df_actt_wrist <- as.data.frame(prop.table(cm_actt_wrist$table, 
                                             margin = 1))
cm_df_actt_wrist$Reference <- factor(cm_df_actt_wrist$Reference, 
                                     levels = rev(levels(cm_df_actt_wrist$Reference)))

plot_actt_wrist <- ggplot(data = cm_df_actt_wrist, aes(x = Prediction, 
                                                       y = Reference, 
                                                       fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.2f", Freq)), 
            color = "white", 
            size = 5) +
  scale_fill_gradient(low = "white", 
                      high = pal[3]) +
  theme(legend.position = "none") +
  labs(title = "ACTT (wrist)", 
       x = "Predicted Class", 
       y = "True Class")

# Composite Figure 3: 2×2 panel, one confusion matrix per device × placement
cowplot::plot_grid(plot_hip, 
                   plot_wrist, 
                   plot_actt_hip, 
                   plot_actt_wrist, 
                   ncol   = 2, 
                   labels = c('a', 'b', 'c', 'd'))
```

# Supplementary analyses

# Effect of sex variable on model performance

```{r}
# ─────────────────────────────────────────────────────────────────────────────
# Performance assessment stratified by sex
# Repeats the classification metrics from the main analysis separately
# for male and female participants to address reviewer concern about
# sex differences in body composition and energy expenditure.
# ─────────────────────────────────────────────────────────────────────────────

# Append sex to dfm from the original df
dfm_sex <- dfm %>%
  dplyr::mutate(fit_MET = fit_MET) %>%
  dplyr::mutate(
    actual_class    = cut(METs,
                          breaks = c(-Inf, 3, 6, 9, Inf),
                          labels = c('[0,3)', '[3,6)', '[6,9)', '[9,inf)'),
                          right  = FALSE),
    predicted_class = calc_met_class(fit_MET)
  )

sex_levels <- dfm_sex %>%
  tidyr::drop_na(Sex) %>%
  dplyr::pull(Sex) %>%
  unique()

# Refit model and append fitted METs
fit_MET_sex <- lm(formula = sqrt(METs) ~ sqrt(act_value) * act,
                  data = dfm_sex) %>%
  fitted() %>%
  `^`(2)

dfm_sex <- dfm_sex %>%
  dplyr::mutate(fit_MET = fit_MET_sex)

# Classify actual and predicted METs
dfm_sex <- dfm_sex %>%
  dplyr::mutate(
    actual_class    = cut(METs,
                          breaks = c(-Inf, 3, 6, 9, Inf),
                          labels = c('[0,3)', '[3,6)', '[6,9)', '[9,inf)'),
                          right  = FALSE),
    predicted_class = calc_met_class(fit_MET)
  )

# --- Confusion matrices per sex ----------------------------------------------
sex_levels <- dfm_sex %>% 
  tidyr::drop_na(Sex) %>% 
  dplyr::pull(Sex) %>% 
  unique()

cm_plots_sex <- purrr::map(sex_levels, function(s) {
  
  d <- dfm_sex %>% dplyr::filter(Sex == s)
  
  cm <- caret::confusionMatrix(d$predicted_class, d$actual_class)
  
  cat('\n===', s, '===\n')
  print(cm)
  
  cm_df <- as.data.frame(prop.table(cm$table, margin = 1))
  cm_df$Reference <- factor(cm_df$Reference,
                            levels = rev(levels(cm_df$Reference)))
  
  ggplot(cm_df, aes(x = Prediction, y = Reference, fill = Freq)) +
    geom_tile() +
    geom_text(aes(label = sprintf('%.2f', Freq)),
              color = 'white', size = 5) +
    scale_fill_gradient(low  = 'white',
                        high = met.brewer('Archambault', 11)[3]) +
    theme(legend.position = 'none') +
    labs(title = s,
         x     = 'Predicted Class',
         y     = 'True Class')
})

cowplot::plot_grid(plotlist = cm_plots_sex,
                  ncol     = 2,
                  labels   = 'AUTO')

# --- AUC per class per sex ---------------------------------------------------
auc_sex <- purrr::map_dfr(sex_levels, function(s) {
  
  d <- dfm_sex %>% dplyr::filter(Sex == s)
  actual <- factor(d$actual_class,
                   levels = c('[0,3)', '[3,6)', '[6,9)', '[9,inf)'))
  
  purrr::map_dfr(levels(actual), function(lvl) {
    binary  <- as.numeric(actual == lvl)
    roc_obj <- pROC::roc(binary, d$fit_MET, quiet = TRUE)
    tibble::tibble(Sex   = s,
                   Class = lvl,
                   AUC   = as.numeric(pROC::auc(roc_obj)))
  })
})

print(auc_sex)
```


## Bland-Altman plots

```{r}
# ─────────────────────────────────────────────────────────────────────────────
# Bland-Altman analysis — predicted METs vs measured METs
# For each device × placement: x = mean(predicted MET, measured MET),
#                               y = predicted MET − measured MET
# ─────────────────────────────────────────────────────────────────────────────

# dfm already contains fit_MET (model-predicted METs) and METs (measured)
# and the act factor with all four device × placement levels
ba_met_data <- dfm %>%
  dplyr::select(ID, Conditions, act, METs, fit_MET) %>%
  dplyr::mutate(
    mean_val = (fit_MET + METs) / 2,
    diff_val = fit_MET - METs,
    device   = factor(act, levels = c('GT3X+ (hip)', 'GT3X+ (wrist)',
                                      'ACTT (hip)',  'ACTT (wrist)'))
  ) %>%
  tidyr::drop_na(mean_val, diff_val)

# Bias and limits of agreement per device × placement
ba_met_stats <- ba_met_data %>%
  dplyr::group_by(device) %>%
  dplyr::summarise(
    bias      = mean(diff_val),
    sd_diff   = sd(diff_val),
    loa_upper = bias + 1.96 * sd_diff,
    loa_lower = bias - 1.96 * sd_diff,
    .groups   = 'drop'
  )

# Per-condition summary: mean ± SD coloured by condition
ba_met_summary <- ba_met_data %>%
  dplyr::group_by(device, Conditions) %>%
  dplyr::summarise(
    mean_x = mean(mean_val),
    mean_y = mean(diff_val),
    sd_y   = sd(diff_val),
    .groups = 'drop'
  )

print(ba_met_stats)

# Bland-Altman plot — predicted vs measured METs
# Bland-Altman plot — predicted vs measured METs
ba_met_data %>%
  dplyr::left_join(ba_met_stats, by = 'device') %>%
  ggplot2::ggplot(aes(x = mean_val, y = diff_val)) +
  # Raw points: no colour or shape mapping
  geom_point(alpha = 0.4, size = 2) +
  # Per-panel reference lines via data + aes
  geom_hline(data = ba_met_stats,
             aes(yintercept = bias),
             colour = pal[4],
             linewidth = 0.9, linetype = 'dashed') +
  geom_hline(data = ba_met_stats,
             aes(yintercept = loa_upper),
             colour = pal[1],
             linewidth = 0.7, linetype = 'dotted') +
  geom_hline(data = ba_met_stats,
             aes(yintercept = loa_lower),
             colour = pal[1],
             linewidth = 0.7, linetype = 'dotted') +
  geom_hline(yintercept = 0,
             colour = 'grey60', linewidth = 0.4) +
  geom_text(
    data = ba_met_stats,
    aes(x = Inf, y = bias,      label = sprintf('Bias: %.2f', bias)),
    hjust = 1.05, vjust = -0.4, size = 3,
    colour = pal[4]
  ) +
  geom_text(
    data = ba_met_stats,
    aes(x = Inf, y = loa_upper, label = sprintf('+1.96 SD: %.2f', loa_upper)),
    hjust = 1.05, vjust = -0.4, size = 3,
    colour = pal[1]
  ) +
  geom_text(
    data = ba_met_stats,
    aes(x = Inf, y = loa_lower, label = sprintf('-1.96 SD: %.2f', loa_lower)),
    hjust = 1.05, vjust = 1.4,  size = 3,
    colour = pal[1]
  ) +
  facet_wrap(~ device, ncol = 2) +
  labs(
    x      = 'Mean of predicted and measured METs',
    y      = 'Difference: predicted − measured METs',
    title  = 'Bland-Altman plots: model-predicted vs measured METs'
  ) +
  theme(legend.position = 'bottom')
```

## Plots test

```{r}
#| warning: false
#| fig-height: 12
#| fig-width: 8

# Panel b — Activity counts, wrist placements only
pb <- df %>% 
  tidyr::pivot_longer(cols = contains(c('GT', 'ACT', 'MET')), 
                      names_to  = 'act', 
                      values_to = 'act_value') %>% 
  dplyr::filter(act != 'METs') %>% 
  dplyr::filter(Conditions != 'Sit-Stands') %>%
  dplyr::filter(stringr::str_detect(act, 'wrist')) %>%
  dplyr::mutate(
    act        = relevel(as.factor(act), ref = 'GT3X+ (wrist)'), 
    Conditions = relevel(as.factor(Conditions), ref = 'Rest'),
    act        = factor(act, levels = c("GT3X+ (wrist)", "ACTT (wrist)"))
  ) %>% 
  ggplot2::ggplot(aes(x = Conditions, y = act_value)) + 
  geom_point(aes(colour = act), 
             position = pos_b,
             size = 2, alpha = 0.4) + 
  geom_text(aes(label = ID, colour = act),
            position = pos_b,
            size = 2, alpha = 0.6,
            vjust = -0.6) +
  geom_boxplot(fill = NA, outliers = FALSE) + 
  geom_smooth(aes(group = act), 
              colour   = 'grey50', 
              linetype = 'dashed', 
              method   = 'loess', 
              se       = FALSE, 
              size     = 0.8) + 
  facet_wrap(~ act, scales = 'free_y', ncol = 1) + 
  scale_colour_manual(values = c(pal[4], pal[3])) + 
  theme(legend.position = "none", 
        axis.text.x     = element_text(angle = 45, hjust = 1)) + 
  ylab(expression("Activity Counts (counts·min"^-1*")")) + 
  xlab(element_blank())
```

