---
title: "ACTT validation study"
format: 
  html:
    self-contained: true
editor_options: 
  chunk_output_type: inline
---

# 📦 Load Required Libraries

```{r}
#| warning: false
library(tidyverse)
library(caret)
library(MuMIn)
library(MetBrewer)
library(broom)
library(cowplot)
library(pROC)
```

# 🎨 Set Theme and Color Palette

```{r}
pal <- met.brewer("Archambault", 11)[c(1,4,6,8)]

theme_set(theme_bw() + 
            theme_bw(base_size = 14, 
                     base_family = "Helvetica") + 
            theme(panel.border = element_rect(linewidth = 1.5)))
```

# 📁 Load and View Data

```{r}
#| warning: false
df <- readr::read_csv(file = 'data/acttrust_validation_full.csv')
df
```

# 🔍 ANOVA and Post-Hoc Tests

## ⚖️️ ANOVA by Device and Condition

```{r}
df %>% 
  dplyr::filter(Conditions != 'Sit-Stands') %>% 
  tidyr::pivot_longer(cols = contains(c('GT', 'ACT')), names_to = 'act', values_to = 'act_value') %>% 
  dplyr::mutate(act = relevel(as.factor(act), ref = 'GT3X+ (hip)')) %>% 
  dplyr::mutate(Conditions = relevel(as.factor(Conditions), ref = 'Rest')) %>% 
  lm(formula = sqrt(act_value) ~ Conditions:act, data = .) %>% 
  aov() %>% 
  summary()
```

## 🔬 Tukey Post-Hoc

```{r}
aov_model <-df %>% 
  dplyr::filter(Conditions != 'Sit-Stands') %>% 
  tidyr::pivot_longer(cols = contains(c('GT', 'ACT')), names_to = 'act', values_to = 'act_value') %>% 
  dplyr::mutate(act = relevel(as.factor(act), ref = 'GT3X+ (hip)')) %>% 
  dplyr::mutate(Conditions = relevel(as.factor(Conditions), ref = 'Rest')) %>% 
  lm(formula = sqrt(act_value) ~ Conditions:act, data = .) %>% 
  aov()

tukey_res <- TukeyHSD(aov_model)
tidy_tukey <- tidy(tukey_res)

tidy_tukey
```

# 🔗 Correlation Matrix of Devices

```{r}
cor_matrix <- df %>% 
  dplyr::filter(Conditions != 'Sit-Stands') %>% 
  dplyr::select(contains(c('GT', 'ACT'))) %>% 
  tidyr::drop_na() %>% 
  cor(., method = "pearson")

cor_melt <- reshape2::melt(cor_matrix)
cor_melt$Var2 <- factor(cor_melt$Var2, levels = rev(levels(cor_melt$Var2)))

ggplot(cor_melt, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient(low = met.brewer("Archambault", 11)[6], 
                      high = met.brewer("Archambault", 11)[4], 
                      name = "Pearson\nCorrelation\n", 
                      labels = scales::label_number(accuracy = 0.01)) +
  xlab(element_blank()) + 
  ylab(element_blank()) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1), 
        axis.text.y = element_text(size = 12),
        legend.title = element_text(size = 14), 
        legend.text = element_text(size = 12)) +
  coord_fixed()
```

# 📈 Linear Model and Predictions

```{r}
dfm <- df %>% 
  dplyr::filter(!Conditions %in% c('Sit-Stands')) %>% 
  dplyr::mutate(Conditions = factor(Conditions)) %>%
  tidyr::drop_na() %>%
  tidyr::pivot_longer(cols = contains(c('GT', 'ACT')), names_to = 'act', values_to = 'act_value') %>% 
  dplyr::mutate(act = relevel(as.factor(act), ref = 'GT3X+ (hip)')) 

lm(formula = sqrt(METs) ~ sqrt(act_value)*act, data = dfm) %>% 
  summary()

```

# 📊 Plot Regression Fits

```{r}

cf <- lm(formula = sqrt(METs) ~ sqrt(act_value)*act, data = dfm) %>% 
  coef()

dfm %>% 
  ggplot2::ggplot(aes(x = sqrt(act_value), y = sqrt(METs))) +
  geom_point(aes(colour = act, shape = Conditions), size = 2.5) + 
  geom_abline(intercept = cf[1], slope = cf[2], colour = pal[1], linewidth = 1.5) +
  geom_abline(intercept = cf[1] + cf[3], slope = cf[2] + cf[6], colour = pal[2], linewidth = 1.5) +
  geom_abline(intercept = cf[1] + cf[4], slope = cf[2] + cf[7], colour = pal[3], linewidth = 1.5) +
  geom_abline(intercept = cf[1] + cf[5], slope = cf[2] + cf[8], colour = pal[4], linewidth = 1.5) +
  scale_colour_manual(values = pal) + 
  guides(colour = guide_legend(override.aes = list(size=5)), 
         shape = guide_legend(override.aes = list(size=5))) + 
  theme(legend.background = element_blank(), 
        legend.title = element_blank()) + 
  xlab(expression(Activity^{1/2})) + 
  ylab(expression(MET^{1/2}))

lm(formula = sqrt(METs) ~ sqrt(act_value)*act, data = dfm) %>% 
  summary()
```

# 🔮 Classification Metrics

## 🤖 Fit and Classify

```{r}
fit_MET <- lm(formula = sqrt(METs) ~ sqrt(act_value)*act, 
              data = dfm)  %>% 
  fitted()

fit_MET <- fit_MET^2
```

```{r}
lm(formula = sqrt(METs) ~ sqrt(act_value)*act, 
              data = dfm) %>% 
  summary()
```


```{r}
# Define the function to classify METs
calc_met_class <- function(met_values) {
  classes <- cut(met_values,
                 breaks = c(-Inf, 3, 6, 9, Inf),
                 labels = c('[0,3)', '[3,6)', '[6,9)', '[9,inf)'),
                 right = FALSE)
  return(classes)
}

dft_pred <- dfm %>% 
  dplyr::select(Conditions, METs, act_value, BMI, ID) %>%
  dplyr::mutate(fit_MET^2) %>% 
  dplyr::mutate(actual_class = cut(METs,
                 breaks = c(-Inf, 3, 6, 9, Inf),
                 labels = c('[0,3)', '[3,6)', '[6,9)', '[9,inf)'),
                 right = FALSE)) %>% 
  dplyr::mutate(predicted_class = calc_met_class(fit_MET))

caret::confusionMatrix(dft_pred$predicted_class, dft_pred$actual_class)
cm <- caret::confusionMatrix(dft_pred$predicted_class, dft_pred$actual_class)
cm_df <- as.data.frame(prop.table(cm$table, margin = 1))

cm_df$Reference <- factor(cm_df$Reference, levels = rev(levels(cm_df$Reference)))
```

## 📊 Confusion Matrix Plot

```{r}
ggplot(data = cm_df, aes(x = Prediction, y = Reference, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.2f", Freq)), color = "white", size = 5) +
  scale_fill_gradient(low = "white", high = met.brewer("Archambault", 11)[3]) +
  theme(legend.position = "none") + 
  labs(title = "Confusion Matrix", x = "Predicted Class", y = "True Class")
```

## 📈 AUC by Class

```{r}
# Prepare
actual_class <- factor(dft_pred$actual_class, 
                       levels = c('[0,3)', '[3,6)', '[6,9)', '[9,inf)'))

# Store results
auc_results <- data.frame(Class = character(), AUC = numeric())

# Loop over each class and compute one-vs-rest AUC
for (lvl in levels(actual_class)) {
  binary_actual <- as.numeric(actual_class == lvl)  # 1 if true class, else 0
  roc_obj <- roc(binary_actual, fit_MET)
  auc_val <- auc(roc_obj)
  auc_results <- rbind(auc_results, data.frame(Class = lvl, AUC = auc_val))
}

print(auc_results)

```


# 🧮 Device-Specific Cutoffs

```{r}
#| warning: false
# Define the cut-off points
cutoffs <- tibble::tibble(MET_values = c(3, 6, 9, 4, 8)) %>% 
  dplyr::mutate(`GT3X+ (hip)` = (((sqrt(MET_values) - cf[1])/cf[2]))^2, 
                `ACTT (hip)` = ((sqrt(MET_values) - cf[1] - cf[3])/(cf[2] + cf[6]))^2, 
                `ACTT (wrist)` = ((sqrt(MET_values) - cf[1] - cf[4])/(cf[2] + cf[7]))^2, 
                `GT3X+ (wrist)` = ((sqrt(MET_values) - cf[1] - cf[5])/(cf[2] + cf[8]))^2) %>% 
  t()

cutoffs[2:5,] %>% 
  tibble::as.tibble() %>% 
  dplyr::mutate(Actimeter = c('GT3X+ (hip)', 'ACTT (hip)', 'ACTT (wrist)', 'GT3X+ (wrist)')) %>% 
  dplyr::rename('3'= V1, '6' = V2, '9' = V3, '4' = V4, '8' = V5) %>%
  gt::gt()
  
```

# 📊 Summary Plots

```{r}
#| warning: false
#| fig-height: 12
#| fig-width: 8

sum_plot <- cowplot::plot_grid(
  df %>% 
  tidyr::pivot_longer(cols = contains(c('GT', 'ACT', 'MET')), names_to = 'act', values_to = 'act_value') %>% 
  dplyr::filter(act == 'METs') %>% 
  dplyr::filter(Conditions != 'Sit-Stands') %>% 
  dplyr::mutate(Conditions = relevel(as.factor(Conditions), ref = 'Rest')) %>% 
  
  ggplot2::ggplot(aes(x = Conditions, y = act_value)) + 
  geom_point(aes(colour = act), position = position_jitterdodge(0.5), size = 3, alpha = 0.4) + 
  geom_boxplot(fill = NA, outliers = FALSE) + 
  geom_smooth(aes(group = act), 
              colour = 'grey50', 
              linetype = 'dashed', 
              method = 'loess', 
              se = FALSE, 
              size = 0.8) + 
  facet_wrap(~act, scales = 'free_y') + 
  scale_colour_manual(values = 'grey70') + 
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1)) + 
  ylab(expression("Metabolic Equivalents (VO"^2*")")) + 
  xlab(element_blank()), 
  df %>% 
  tidyr::pivot_longer(cols = contains(c('GT', 'ACT', 'MET')), names_to = 'act', values_to = 'act_value') %>% 
  dplyr::filter(act != 'METs') %>% 
  dplyr::filter(Conditions != 'Sit-Stands') %>% 
  dplyr::mutate(act = relevel(as.factor(act), ref = 'GT3X+ (hip)'), 
                Conditions = relevel(as.factor(Conditions), ref = 'Rest')) %>% 
  dplyr::mutate(act = factor(act, levels = c("GT3X+ (hip)", "GT3X+ (wrist)", "ACTT (hip)", "ACTT (wrist)"))) %>% 

  
  ggplot2::ggplot(aes(x = Conditions, y = act_value)) + 
  geom_point(aes(colour = act), position = position_jitterdodge(1.2), size = 2, alpha = 0.4) + 
  geom_boxplot(fill = NA, outliers = FALSE) + 
  geom_smooth(aes(group = act), 
              colour = 'grey50', 
              linetype = 'dashed', 
              method = 'loess', 
              se = FALSE, 
              size = 0.8) + 
  facet_wrap(~act, scales = 'free_y') + 
  scale_colour_manual(values = c(pal[1],pal[4],pal[2],pal[3])) + 
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1)) + 
  ylab(expression("Activity Counts (counts·min"^-1*")")) + 
  xlab(element_blank()), 
  ncol = 1, 
  labels = c('a', 'b'))
sum_plot
```

# Device-specific confusion matrix

```{r}
# First, calculate fit_MET and add it to dfm
fit_MET <- lm(formula = sqrt(METs) ~ sqrt(act_value) * act, data = dfm) %>%
  fitted() %>%
  `^`(2)  # Square the fitted values

# Append fit_MET as a column
dfm <- dfm %>%
  mutate(fit_MET = fit_MET)

# Define device types
device_types <- c('GT3X+ (hip)', 'GT3X+ (wrist)', 'ACTT (hip)', 'ACTT (wrist)')

# Store plots
plots <- list()

# GT3X+ (hip)
dft_hip <- dfm %>%
  filter(act == 'GT3X+ (hip)') %>%
  select(Conditions, METs, act_value, BMI, ID, fit_MET) %>%
  mutate(
    actual_class = cut(METs, breaks = c(-Inf, 3, 6, 9, Inf),
                       labels = c('[0,3)', '[3,6)', '[6,9)', '[9,inf)'), right = FALSE),
    predicted_class = calc_met_class(fit_MET)
  )

cm_hip <- caret::confusionMatrix(dft_hip$predicted_class, dft_hip$actual_class)
cm_df_hip <- as.data.frame(prop.table(cm_hip$table, margin = 1))
cm_df_hip$Reference <- factor(cm_df_hip$Reference, levels = rev(levels(cm_df_hip$Reference)))

plot_hip <- ggplot(data = cm_df_hip, aes(x = Prediction, y = Reference, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.2f", Freq)), color = "white", size = 5) +
  scale_fill_gradient(low = "white", high = pal[1]) +
  theme(legend.position = "none") +
  labs(title = "GT3X+ (hip)", x = "Predicted Class", y = "True Class")


# GT3X+ (wrist)
dft_wrist <- dfm %>%
  filter(act == 'GT3X+ (wrist)') %>%
  select(Conditions, METs, act_value, BMI, ID, fit_MET) %>%
  mutate(
    actual_class = cut(METs, breaks = c(-Inf, 3, 6, 9, Inf),
                       labels = c('[0,3)', '[3,6)', '[6,9)', '[9,inf)'), right = FALSE),
    predicted_class = calc_met_class(fit_MET)
  )

cm_wrist <- caret::confusionMatrix(dft_wrist$predicted_class, dft_wrist$actual_class)
cm_df_wrist <- as.data.frame(prop.table(cm_wrist$table, margin = 1))
cm_df_wrist$Reference <- factor(cm_df_wrist$Reference, levels = rev(levels(cm_df_wrist$Reference)))

plot_wrist <- ggplot(data = cm_df_wrist, aes(x = Prediction, y = Reference, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.2f", Freq)), color = "white", size = 5) +
  scale_fill_gradient(low = "white", high = pal[4]) +
  theme(legend.position = "none") +
  labs(title = "GT3X+ (wrist)", x = "Predicted Class", y = "True Class")


# ACTT (hip)
dft_actt_hip <- dfm %>%
  filter(act == 'ACTT (hip)') %>%
  select(Conditions, METs, act_value, BMI, ID, fit_MET) %>%
  mutate(
    actual_class = cut(METs, breaks = c(-Inf, 3, 6, 9, Inf),
                       labels = c('[0,3)', '[3,6)', '[6,9)', '[9,inf)'), right = FALSE),
    predicted_class = calc_met_class(fit_MET)
  )

cm_actt_hip <- caret::confusionMatrix(dft_actt_hip$predicted_class, dft_actt_hip$actual_class)
cm_df_actt_hip <- as.data.frame(prop.table(cm_actt_hip$table, margin = 1))
cm_df_actt_hip$Reference <- factor(cm_df_actt_hip$Reference, levels = rev(levels(cm_df_actt_hip$Reference)))

plot_actt_hip <- ggplot(data = cm_df_actt_hip, aes(x = Prediction, y = Reference, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.2f", Freq)), color = "white", size = 5) +
  scale_fill_gradient(low = "white", high = pal[2]) +
  theme(legend.position = "none") +
  labs(title = "ACTT (hip)", x = "Predicted Class", y = "True Class")


# ACTT (wrist)
dft_actt_wrist <- dfm %>%
  filter(act == 'ACTT (wrist)') %>%
  select(Conditions, METs, act_value, BMI, ID, fit_MET) %>%
  mutate(
    actual_class = cut(METs, breaks = c(-Inf, 3, 6, 9, Inf),
                       labels = c('[0,3)', '[3,6)', '[6,9)', '[9,inf)'), right = FALSE),
    predicted_class = calc_met_class(fit_MET)
  )

cm_actt_wrist <- caret::confusionMatrix(dft_actt_wrist$predicted_class, dft_actt_wrist$actual_class)
cm_df_actt_wrist <- as.data.frame(prop.table(cm_actt_wrist$table, margin = 1))
cm_df_actt_wrist$Reference <- factor(cm_df_actt_wrist$Reference, levels = rev(levels(cm_df_actt_wrist$Reference)))

plot_actt_wrist <- ggplot(data = cm_df_actt_wrist, aes(x = Prediction, y = Reference, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.2f", Freq)), color = "white", size = 5) +
  scale_fill_gradient(low = "white", high = pal[3]) +
  theme(legend.position = "none") +
  labs(title = "ACTT (wrist)", x = "Predicted Class", y = "True Class")

cowplot::plot_grid(plot_hip, 
                   plot_wrist, 
                   plot_actt_hip, 
                   plot_actt_wrist, 
                   ncol = 2, 
                   labels = c('a', 'b', 'c', 'd'))
```
